Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model

### 2013-12-09 Paper/gammaRays.tex
Drafted 1/3 of introduction

Gamma ray bursts are the most energetic explosions in the Universe, therefore they are of considerable interest for particle physics. Recently developed instruments, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{SwiftReference} allowed for extensive study of gamma ray bursts \cite{Vianello:2013ela, Gehrels:2013xd}. Many discoveries were made: for example, it was found that the energy emitted by GRBs in gamma rays remains relatively constant for different bursts \cite{Bloom:2003wy}. Temporal variations of GRB spectra were studied. The spectral lags were observed between different low-energy bands \cite{Yi:2005ht}, as well as between low-energy and high energy radiation \cite{}. However, recent observations of luminous bursts by Fermi allowed to study temporal variations even in high energy spectra. In this study we used KS-test to search for spectral lags between two high energy bands: $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.

>>>>

Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics. Recent observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} allowed for extensive study of these explosions \cite{Vianello:2013ela,Gehrels:2013xd}. These studies led to several interesting results. For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}. Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}. The very high energy radiation was discovered to be extended if compared to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}. Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$. In this paper we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
